\bf Adapting the Chumbley Score to match striae on Land Engraved Areas of bullets

### 2018-08-18 doc/Response.Rmd

*Thank you for your comments, we appreciate your insight. You are right, for the purpose of matching bullet to bullet, all lands of each of the bullets and their relative positions on the bullet should be considered. This is not the problem we tried to address in this manuscript. To get from land-to-land comparisons to bullet-to-bullet comparisons, an additional step is required. In \cite{chu2010}, the average correlation value is taken. In this manuscript, the land-to-land comparisons result in a p-value. There are different options to aggregate these values at this stage to quantify a bullet-bullet comparison.
*

*We agree on the point that the toolmarks have large differences in width and the parameters of the chumbley algorithm are optimized for tools, which is why we first do a parameter study to identify the best set of parameters that would work for the LEAs.*

Another very important point;  in none of the cited texts should there be an impression that if a ground truth match is not indicated by sufficient match score, that it falls into a miss-identification, or Type 2 error "a failed test".  A type 2 error (false positive) would occur if a comparison has a match score over a threshold for identification when the ground truth is two sources.  Therefore, the hot-button term "error rates" in this case is not accurate.
The manuscript is using this erroneous description throughout, is other words, "Missed identifications" and are improperly indicated as described as "miss-identifications".  This miss-characterization is a major weakness in the manuscript.

*You are right - the term 'error rate' is highly charged emotionally for firearms examiners. The Chumbley score test is peculiar in the sense, that the null hypothesis is stated as 'two markings coming from a different source'. For clarification, we have added a paragraph at the end of the section on the the Chumbley score that details the test hypotheses with the corresponding (statistical) errors.
We have also added a sentence to the effect that all of the comparisons run on the Hamby 252 set only included lands that were deemed suitable for a comparison (e.g. not affected by tank rash), similar to Hare et al (2016). Lands leading to inconclusive results by FTEs are therefore not included in the comparisons and would not lead to a 'missed identifaction'.
*

### 2018-08-18 doc/Response.Rmd

*Thank you for your comments, we appreciate your insight. You are right, for the purpose of matching bullet to bullet, all lands of each of the bullets and their relative positions on the bullet should be considered. This is not the problem we tried to address in this manuscript. To get from land-to-land comparisons to bullet-to-bullet comparisons, an additional step is required. In \cite{chu2010}, the average correlation value is taken. In this manuscript, the land-to-land comparisons result in a p-value. There are different options to aggregate these values at this stage to quantify a bullet-bullet comparison.
*

*We agree on the point that the toolmarks have large differences in width and the parameters of the chumbley algorithm are optimized for tools, which is why we first do a parameter study to identify the best set of parameters that would work for the LEAs.*

Another very important point;  in none of the cited texts should there be an impression that if a ground truth match is not indicated by sufficient match score, that it falls into a miss-identification, or Type 2 error "a failed test".  A type 2 error (false positive) would occur if a comparison has a match score over a threshold for identification when the ground truth is two sources.  Therefore, the hot-button term "error rates" in this case is not accurate.
The manuscript is using this erroneous description throughout, is other words, "Missed identifications" and are improperly indicated as described as "miss-identifications".  This miss-characterization is a major weakness in the manuscript.

*You are right - the term 'error rate' is highly charged emotionally for firearms examiners. The Chumbley score test is peculiar in the sense, that the null hypothesis is stated as 'two markings coming from a different source'. For clarification, we have added a paragraph at the end of the section on the the Chumbley score that details the test hypotheses and a paragraph at the end of the following section with the corresponding (statistical) errors.
We have also added a sentence to the effect that all of the comparisons run on the Hamby 252 set only included lands that were deemed suitable for a comparison (e.g. not affected by tank rash), similar to Hare et al (2016). Lands leading to inconclusive results by FTEs are therefore not included in the comparisons and would not lead to a 'missed identifaction'.
*

@@

Coarseness of the surface is not a direct indicator of sub-class influence.  More literature review is necessary to understand the earmarks of subclass.  For the surfaces compared in the Hamby sets, there was no indication of subclass influence between fired bullet sets from consecutively rifled barrels.  If the term is instead a parameter of form (curvature) then that should be more fully described, and how it is used. (I suspect that this is how it was used for fired bullets in this research.)

*A coarse Gaussian filter was indeed used to remove some class characteristics in Bachrach et al (2010) for toolmark striae. In this manuscript, the filter is used to remove bullet curvature. However, subclass characteristics expressed in form of 'waviness' in the barrel (e.g. introduced by specifics in the firing process) can be addressed in the same way as for toolmarks.*

@@

Additionally, consider including a firearm examiner who is acquainted with objective methods of comparisons for 3D toolmark profiles to review your updated manuscript draft.

*Our research center is working closely with forensic examiners, including firearms examiners. Thank you for the advice.*

### 2018-08-18 doc/paper.Rmd
hypotheses  and error rates

---
title: "Adapting the Chumbley Score to match striae on Land Engraved Areas of bullets"
authors:
# - affiliation: Department of Statistics, Iowa State University 
- name: Ganesh Krishnan,
  degree: Master of Science (M.Sc.)
  # thank: The authors gratefully acknowledge funding from CSAFE through cooperative agreement \# 70NANB15H176 between NIST and Iowa State University. We would also like to thank the efforts of CSAFE and Roy J carver Microscopy lab in scanning the Hamby-44 set and providing the scans to us.
# - affiliation: Department of Statistics and CSAFE, Iowa State University 
- name: Prof. Heike Hofmann,
  degree: Doctor of Philosophy (Ph.D.)
affiliation: Department of Statistics, Iowa State University and CSAFE (Center for Statistics and Applications in Forensic Evidence)
funding: The authors gratefully acknowledge funding from CSAFE through cooperative agreement \# 70NANB15H176 between NIST and Iowa State University. 
thank: We would like to thank the efforts of CSAFE and Roy J Carver Microscopy lab in scanning the Hamby-44 set and providing the scans to us. We would also like to thank NIST for providing the Hamby-44 bullets to us for scanning.
presentedat:  This work was presented in the form of a poster in the AFTE 2018 49th Annual Training Seminar, held from June 3rd to June 8th, 2018 in Charleston, West Virginia
biblio-style: jfs
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    keep_tex: yes
    latex_engine: pdflatex
    template: template.tex
  html_document: default
blinded: 0
keywords:
- forensic science
- toolmark
- cross-correlation
- Mann-Whitney U statistic
- land engraved areas (LEAs)
- algorithm
- signatures 
- same-source problem
bibliography: bibliography
abstract: The same-source problem remains a major challenge in forensic toolmark and firearm examination. Technological advances in surface imaging allow measurements of 3D surfaces at previously unforeseen resolutions and enable digitized imaging. Here, we investigate the applicability of the Chumbley scoring method \cite{hadler}, developed for screwdriver markings, for same-source identification of bullet striae. We provide methods to identify parameters that minimize the error rates for matching of LEAs using Hamby datasets 44 and 252 measured by NIST and CSAFE. We suggest a remedial algorithm to alleviate the problem of failed tests in the method, increasing both the power of the test and reducing error rates. Type II error rates of the proposed method improve by more than one third (Type I error of 0.05) and are on average about 0.22. This puts the proposed method on similar footing as other single feature matching approaches in the literature. 
---

@@

\begin{table}[ht]
\centering
\resizebox{\textwidth}{!}{\begin{tabular}{rlrrr}
  \hline
Method & Data Source & False Positives & False Negatives\\
  \hline
  \textbf{\citet{aoas}} & Hamby 252 (NIST scans) \\
Consecutive Matching Bullets Striae (CMS) &   &  6.25\% &  33.85\% \\ 
Consecutive Non-matching Striae (CNMS)  &  &  6.26\% & 35.42\%  \\ 
Average Distance (D) & Winchester  &   6.25\% & 45.83\% \\ 
Cross-correlation Function (CCF) & &  6.25\% & 17.71\%  \\ 
Sum of Peaks (S) &  &  6.25\% &  18.23\% \\
   \hline
   \textbf{\citet{chu2013}} & Hamby 252 (NIST scans) \\
Consecutive Matching Bullets Striae (CMS) &  &  0\% &  52\% \\ 
   \hline
   \textbf{\citet{chu2010}} & \small{6 types of firearms, 2 types of ammunition} \\
   \textbf{\citet{ma2004}} & \small{NIST standard bullet SRM 8240} \\
Cross-Correlation Function (CCF) &   &  - &  - \\ 
    \hline
\end{tabular}}
\caption{\label{tab:bullets-ER} Error Rates in same-source single feature land-to-land analysis reported in the literature. \gk{Note that differences in error rates do not only come from different methods but also differently marking firearms and types of ammunition}}
\end{table}

@@

\begin{figure}
\centering
\includegraphics[width=\textwidth]{images/B6-B2-L6-rescaled.png}
```{r, eval=FALSE, echo=FALSE, fig.width=8, fig.height=3, out.width='\\textwidth'}
land <- read.csv("../data/b6b2l6.csv")
land %>% ggplot(aes(x = y, y=resid)) +geom_line() +theme_bw() +
  ylab(expression(paste("Depth (in ",mu,m,")", sep=""))) + 
  xlab(expression(paste("Relative Location (in ",mu,m,")", sep="")))
```

@@

\gk{Bullet striae are most pronounced at the bottom of the bullet heel (because the heel has the most contact with the insides of the barrel). Howver, these areas are also affected by the break off due to friction effects between barrel and the bullet. An optimal crosscut is chosen orthogonally to the striae, close to the heel while avoiding break off, see also \citep{aoas}}. Comparisons of striae from bullets are usually based on comparisons of striae in land engraved areas, which are extracted in form of cross sections, called *profiles* \citep{aoas,ma2004}. From profiles bullet *signatures* \citep{chu2013,aoas} are extracted as residuals of a loess fit or Gaussian filter. This effectively removes topographic structure from the data in the attempt to increase the signal to noise ratio. \autoref{fig:rgl}  shows how the signature from a bullet land (bottom) lines up with the image of the land (top) from which it was extracted. We can see in the figure how the depth and relative position of the striation markings seen in the image are interpreted as peaks and valleys in the signature. <!--The span of the loess fit was found using cross-validation, as described by \citet{aoas}.-->

@@

For both same- and different-shift pairs correlations between the  markings are calculated. 
The intuition here is that for two markings from the same-source the correlation for the in-phase sample should be high, while the correlations of the out-of-phase sample  provide a measure for the base-level correlation for non-matching marks of a given length $w_v$. The Chumbley score is then computed as a Mann Whitney U statistic to compare between in-phase sample and out-of-phase sample. 
In the original method proposed in \citet{chumbley} both in-phase and out-of-phase sample are extracted randomly, whereas \citet{hadler} proposed the above specified deterministic rules for both samples to make the resulting score deterministic while simultaneously avoiding overlaps within selected marks to ensure independence.

@@

Looking closer at \autoref{eq.start}, we see that by definition, some number of tests will fail to produce a result. The problem of failed tests is first mentioned in \citet{afte-chumbley}. Unfortunately, the authors do not provide any percentage of how many tests failed for their data.
The algorithm fails to produce for two reasons: either the number of eligible same-shift pairs is zero, or the number of different-shift pairs is zero.

@@

We then define the pairs for different-shifts as
\begin{equation}\label{eq.diff2}
\left(s_{(j)}^{(1)}, s_{(I-j+1)}^{(1)} \right) \text{ for } j = 
\begin{cases}
1, ..., I & \text{ for even } I \\
1, ..., (I-1)/2, (I-1)/2 + 2, ..., I & \text{ for odd } I
\end{cases},
\end{equation}
i.e.\ for an odd number of same-shift correlations, we skip the middle pair for the different-shift correlations (see \autoref{sketch-diff-2}).
This pairing ensures that the number of different-shift pairings is the same or at most one less than the number of same-shift pairings in all tests.
In the remainder of the paper, we will refer to the algorithm defined by \citet{hadler} as \textbf{(CS1)} and the suggested modified algorithm as \textbf{(CS2)} and compare their performance on the available scans of the Hamby study. The section \ref{appendix:appxfailed} of the Appendix discusses about the various scenarious of failed tests.
\begin{figure}[hbtp]
\centering
\includegraphics[width=.7\textwidth]{images/sketch-diff-2.png}
\caption{\label{sketch-diff-2}Sketch of adjusted different-shift pairings. At most one of the same-shift pairings can not be matched with a different-shift pair under algorithm (CS2). }
\end{figure}

>>>>

---
title: "Adapting the Chumbley Score to match striae on Land Engraved Areas of bullets"
affiliation: Department of Statistics, Iowa State University and CSAFE (Center for
  Statistics and Applications in Forensic Evidence)
authors:
- degree: Master of Science (M.Sc.)
  name: Ganesh Krishnan,
- degree: Doctor of Philosophy (Ph.D.)
  name: Prof. Heike Hofmann,
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    keep_tex: yes
    latex_engine: pdflatex
    template: template.tex
  html_document: default
bibliography: bibliography
blinded: 0
funding: The authors gratefully acknowledge funding from CSAFE through cooperative
  agreement \# 70NANB15H176 between NIST and Iowa State University.
keywords:
- forensic science
- toolmark
- cross-correlation
- Mann-Whitney U statistic
- land engraved areas (LEAs)
- algorithm
- signatures
- same-source problem
biblio-style: jfs
presentedat: This work was presented in the form of a poster in the AFTE 2018 49th
  Annual Training Seminar, held from June 3rd to June 8th, 2018 in Charleston, West
  Virginia
thank: We would like to thank the efforts of CSAFE and Roy J Carver Microscopy lab
  in scanning the Hamby-44 set and providing the scans to us. We would also like to
  thank NIST for providing the Hamby-44 bullets to us for scanning.
abstract: The same-source problem remains a major challenge in forensic toolmark and
  firearm examination. Technological advances in surface imaging allow measurements
  of 3D surfaces at previously unforeseen resolutions and enable digitized imaging.
  Here, we investigate the applicability of the Chumbley scoring method \cite{hadler},
  developed for screwdriver markings, for same-source identification of bullet striae.
  We provide methods to identify parameters that minimize the error rates for matching
  of LEAs using Hamby datasets 44 and 252 measured by NIST and CSAFE. We suggest a
  remedial algorithm to alleviate the problem of failed tests in the method, increasing
  both the power of the test and reducing error rates. Type II error rates of the
  proposed method improve by more than one third (Type I error of 0.05) and are on
  average about 0.22. This puts the proposed method on similar footing as other single
  feature matching approaches in the literature.
---

@@

\begin{table}[ht]
\centering
\resizebox{\textwidth}{!}{\begin{tabular}{rlrrr}
  \hline
Method & Data Source & False Positives & False Negatives\\
  \hline
  \textbf{\citet{aoas}} & Hamby 252 (NIST scans) \\
Consecutive Matching Bullets Striae (CMS) &   &  6.25\% &  33.85\% \\ 
Consecutive Non-matching Striae (CNMS)  &  &  6.26\% & 35.42\%  \\ 
Average Distance (D) &   &   6.25\% & 45.83\% \\ 
Cross-correlation Function (CCF) & &  6.25\% & 17.71\%  \\ 
Sum of Peaks (S) &  &  6.25\% &  18.23\% \\
   \hline
   \textbf{\citet{chu2013}} & Hamby 252 (NIST scans) \\
Consecutive Matching Bullets Striae (CMS) &  &  0\% &  52\% \\ 
   \hline
   \textbf{\citet{chu2010}} & \small{6 types of firearms, 2 types of ammunition} \\
   \textbf{\citet{ma2004}} & \small{NIST standard bullet SRM 8240} \\
Cross-Correlation Function (CCF) &   &  - &  - \\ 
    \hline
\end{tabular}}
\caption{\label{tab:bullets-ER} Error Rates in same-source single feature land-to-land analysis reported in the literature. \hh{Note that only error rates from the same data source can be compared directly between methods. Error rates from different sources are influenced by type of firearm and ammunition used.}}
\end{table}

@@

\begin{figure}
\centering
\includegraphics[width=\textwidth]{images/B6-B2-L6-rescaled.pdf}
```{r, eval=FALSE, echo=FALSE, fig.width=8, fig.height=3, out.width='\\textwidth'}
land <- read.csv("../data/b6b2l6.csv")
land %>% ggplot(aes(x = y, y=resid)) +geom_line() +theme_bw() +
  ylab(expression(paste("Depth (in ",mu,m,")", sep=""))) + 
  xlab(expression(paste("Relative Location (in ",mu,m,")", sep="")))
```

@@

Comparisons of striae from bullets are usually based on comparisons of striae in land engraved areas, which are extracted in form of cross sections, called *profiles* \citep{aoas,ma2004}.
Bullet striae are most pronounced at the bottom of the bullet heel (because the heel has the most contact with the insides of the barrel). Howver, these areas are also affected by the break off due to friction effects between barrel and the bullet. An optimal cross section is chosen orthogonally to the striae, close to the heel while avoiding break off as shown in \autoref{fig:rgl}, see also \citep{aoas} for mathematical details of the extraction.

 From profiles bullet *signatures* \citep{chu2013,aoas} are extracted as residuals of a loess fit or Gaussian filter. This effectively removes topographic structure from the data in the attempt to increase the signal to noise ratio. \autoref{fig:rgl}  shows how the signature from a bullet land (bottom) lines up with the image of the land (top) from which it was extracted. We can see in the figure how the depth and relative position of the striation markings seen in the image are interpreted as peaks and valleys in the signature. <!--The span of the loess fit was found using cross-validation, as described by \citet{aoas}.-->

@@

For both same- and different-shift pairs correlations between the  markings are calculated. 
The intuition here is that for two markings from the same-source the correlation for the in-phase sample should be high, while the correlations of the out-of-phase sample  provide a measure for the base-level correlation for non-matching marks of a given length $w_v$.
\hh{More specifically, the null hypothesis of the Chumbley score test is stated as **$H_o$: the markings come from two different sources**} with the alternative given as **$H_a$: the markings come from the same source**. A $p$-value for the Chumbley score test is } then computed as a Mann Whitney U statistic to compare between in-phase sample and out-of-phase sample. A low $p$-value is then interpreted as a rejection of the null hypothesis in favor of the alternative hypothesis of `same-source`.}

In the original method proposed in \citet{chumbley} both in-phase and out-of-phase sample are extracted randomly, whereas \citet{hadler} proposed the above specified deterministic rules for both samples to make the resulting score deterministic while simultaneously avoiding overlaps within selected marks to ensure independence.

@@

Looking closer at \autoref{eq.start}, we see that by definition, some number of tests will fail to produce a result. \hh{Note that this problem is different from erroneous test results.} The problem of failed tests is first mentioned in \citet{afte-chumbley}. Unfortunately, the authors do not provide any percentage of how many tests failed for their data. 
The algorithm fails to produce for two reasons: either the number of eligible same-shift pairs is zero, or the number of different-shift pairs is zero.

@@

We then define the pairs for different-shifts as
\begin{equation}\label{eq.diff2}
\left(s_{(j)}^{(1)}, s_{(I-j+1)}^{(1)} \right) \text{ for } j = 
\begin{cases}
1, ..., I & \text{ for even } I \\
1, ..., (I-1)/2, (I-1)/2 + 2, ..., I & \text{ for odd } I
\end{cases},
\end{equation}
i.e.\ for an odd number of same-shift correlations, we skip the middle pair for the different-shift correlations (see \autoref{sketch-diff-2}).
This pairing ensures that the number of different-shift pairings is the same or at most one less than the number of same-shift pairings in all tests.
In the remainder of the paper, we will refer to the algorithm defined by \citet{hadler} as \textbf{(CS1)} and the suggested modified algorithm as \textbf{(CS2)} and compare their performance on the available scans of the Hamby study.

\hh{Performance of tests is measured with respect to the errors they make in situations where ground truth is known. We distinguish between two error rates: (1) false positive and (2) false negative rate.
*False positives* (or *false identifications*) are situations where the test indicates a match (i.e. the test falsely rejects) but the markings come from different sources. This is also known as a *Type I* error. 
*False negatives* (or *missed identifications*) are situations where the test fails to reject, i.e. the test indicates that the markings come from different sources, but in fact the markings come from the same source. This is a *Type II* error.}

\hh{Note that in all of the following land-to-land comparisons only lands are compared that are suitable for a comparison, i.e. a signature can be extracted from the scan. In particular, lands which exhibited tank rash  were removed from comparison \citep{aoas}.
}

The section \ref{appendix:appxfailed} of the Appendix discusses about the various scenarious of failed tests.
\begin{figure}[hbtp]
\centering
\includegraphics[width=.7\textwidth]{images/sketch-diff-2.png}
\caption{\label{sketch-diff-2}Sketch of adjusted different-shift pairings. At most one of the same-shift pairings can not be matched with a different-shift pair under algorithm (CS2). }
\end{figure}

### 2018-08-18 doc/paper.tex
hypotheses  and error rates

{\textcolor{blue}{changed}} These digitized markings allow the use of
statistical methods which can quantify the scientific mechanism of
comparing markings and serve as basis for an error rate calculation. In
\cite{manytoolmarks2} a relative distance metric is defined and used as
similarity measure between two toolmarks. This approach is expanded in
\citet{manytoolmarks1}: a set of small segments in the markings of two
toolmarks are extracted and similarity is compared using a maximum
Pearson correlation coefficient. The Chumbley scoring method, first
introduced in \cite{chumbley}, uses a similar but more extensive
framework based on a Mann-Whitney U test of the resulting correlation
coefficients. This approach is non-deterministic, because segments are
chosen randomly. In \cite{hadler} the score is fixed to be deterministic
for each pair of toolmarks by choosing segments for comparison
systematically. This approach also ensures independence between segments
of striae.

@@

\begin{table}[ht]
\centering
\resizebox{\textwidth}{!}{\begin{tabular}{rlrrr}
  \hline
Method & Data Source & False Positives & False Negatives\\
  \hline
  \textbf{\citet{aoas}} & Hamby 252 (NIST scans) \\
Consecutive Matching Bullets Striae (CMS) &   &  6.25\% &  33.85\% \\ 
Consecutive Non-matching Striae (CNMS)  &  &  6.26\% & 35.42\%  \\ 
Average Distance (D) & Winchester  &   6.25\% & 45.83\% \\ 
Cross-correlation Function (CCF) & &  6.25\% & 17.71\%  \\ 
Sum of Peaks (S) &  &  6.25\% &  18.23\% \\
   \hline
   \textbf{\citet{chu2013}} & Hamby 252 (NIST scans) \\
Consecutive Matching Bullets Striae (CMS) &  &  0\% &  52\% \\ 
   \hline
   \textbf{\citet{chu2010}} & \small{6 types of firearms, 2 types of ammunition} \\
   \textbf{\citet{ma2004}} & \small{NIST standard bullet SRM 8240} \\
Cross-Correlation Function (CCF) &   &  - &  - \\ 
    \hline
\end{tabular}}
\caption{\label{tab:bullets-ER} Error Rates in same-source single feature land-to-land analysis reported in the literature. {\textcolor{blue}{Note that differences in error rates do not only come from different methods but also differently marking firearms and types of ammunition}}}
\end{table}

@@

\begin{figure}
\centering
\includegraphics[width=\textwidth]{images/B6-B2-L6-rescaled.png}

@@

{\textcolor{blue}{Bullet striae are most pronounced at the bottom of the bullet heel (because the heel has the most contact with the insides of the barrel). Howver, these areas are also affected by the break off due to friction effects between barrel and the bullet. An optimal crosscut is chosen orthogonally to the striae, close to the heel while avoiding break off, see also \citep{aoas}}}.
Comparisons of striae from bullets are usually based on comparisons of
striae in land engraved areas, which are extracted in form of cross
sections, called \emph{profiles} \citep{aoas,ma2004}. From profiles
bullet \emph{signatures} \citep{chu2013,aoas} are extracted as residuals
of a loess fit or Gaussian filter. This effectively removes topographic
structure from the data in the attempt to increase the signal to noise
ratio. \autoref{fig:rgl} shows how the signature from a bullet land
(bottom) lines up with the image of the land (top) from which it was
extracted. We can see in the figure how the depth and relative position
of the striation markings seen in the image are interpreted as peaks and
valleys in the signature.

@@

For both same- and different-shift pairs correlations between the
markings are calculated. The intuition here is that for two markings
from the same-source the correlation for the in-phase sample should be
high, while the correlations of the out-of-phase sample provide a
measure for the base-level correlation for non-matching marks of a given
length \(w_v\). The Chumbley score is then computed as a Mann Whitney U
statistic to compare between in-phase sample and out-of-phase sample. In
the original method proposed in \citet{chumbley} both in-phase and
out-of-phase sample are extracted randomly, whereas \citet{hadler}
proposed the above specified deterministic rules for both samples to
make the resulting score deterministic while simultaneously avoiding
overlaps within selected marks to ensure independence.

@@

Looking closer at \autoref{eq.start}, we see that by definition, some
number of tests will fail to produce a result. The problem of failed
tests is first mentioned in \citet{afte-chumbley}. Unfortunately, the
authors do not provide any percentage of how many tests failed for their
data. The algorithm fails to produce for two reasons: either the number
of eligible same-shift pairs is zero, or the number of different-shift
pairs is zero.

@@

We then define the pairs for different-shifts as
\begin{equation}\label{eq.diff2}
\left(s_{(j)}^{(1)}, s_{(I-j+1)}^{(1)} \right) \text{ for } j = 
\begin{cases}
1, ..., I & \text{ for even } I \\
1, ..., (I-1)/2, (I-1)/2 + 2, ..., I & \text{ for odd } I
\end{cases},
\end{equation} i.e.~for an odd number of same-shift correlations, we
skip the middle pair for the different-shift correlations (see
\autoref{sketch-diff-2}). This pairing ensures that the number of
different-shift pairings is the same or at most one less than the number
of same-shift pairings in all tests. In the remainder of the paper, we
will refer to the algorithm defined by \citet{hadler} as \textbf{(CS1)}
and the suggested modified algorithm as \textbf{(CS2)} and compare their
performance on the available scans of the Hamby study. The section
\ref{appendix:appxfailed} of the Appendix discusses about the various
scenarious of failed tests.

@@

>>>>

{\textcolor{blue}{changed}} These digitized markings allow the use of
statistical methods to quantify the scientific mechanism of comparing
markings and serve as basis for an error rate calculation. In
\cite{manytoolmarks2} a relative distance metric is defined and used as
similarity measure between two toolmarks. This approach is expanded in
\citet{manytoolmarks1}: a set of small segments in the markings of two
toolmarks are extracted and similarity is compared using a maximum
Pearson correlation coefficient. The Chumbley scoring method, first
introduced in \cite{chumbley}, uses a similar but more extensive
framework based on a Mann-Whitney U test of the resulting correlation
coefficients. This approach is non-deterministic, because segments are
chosen randomly. In \cite{hadler} the score is fixed to be deterministic
for each pair of toolmarks by choosing segments for comparison
systematically. This approach also ensures independence between segments
of striae.

@@

\begin{table}[ht]
\centering
\resizebox{\textwidth}{!}{\begin{tabular}{rlrrr}
  \hline
Method & Data Source & False Positives & False Negatives\\
  \hline
  \textbf{\citet{aoas}} & Hamby 252 (NIST scans) \\
Consecutive Matching Bullets Striae (CMS) &   &  6.25\% &  33.85\% \\ 
Consecutive Non-matching Striae (CNMS)  &  &  6.26\% & 35.42\%  \\ 
Average Distance (D) &   &   6.25\% & 45.83\% \\ 
Cross-correlation Function (CCF) & &  6.25\% & 17.71\%  \\ 
Sum of Peaks (S) &  &  6.25\% &  18.23\% \\
   \hline
   \textbf{\citet{chu2013}} & Hamby 252 (NIST scans) \\
Consecutive Matching Bullets Striae (CMS) &  &  0\% &  52\% \\ 
   \hline
   \textbf{\citet{chu2010}} & \small{6 types of firearms, 2 types of ammunition} \\
   \textbf{\citet{ma2004}} & \small{NIST standard bullet SRM 8240} \\
Cross-Correlation Function (CCF) &   &  - &  - \\ 
    \hline
\end{tabular}}
\caption{\label{tab:bullets-ER} Error Rates in same-source single feature land-to-land analysis reported in the literature. {\textcolor{orange}{Note that only error rates from the same data source can be compared directly between methods. Error rates from different sources are influenced by type of firearm and ammunition used.}}}
\end{table}

@@

\begin{figure}
\centering
\includegraphics[width=\textwidth]{images/B6-B2-L6-rescaled.pdf}

@@

Comparisons of striae from bullets are usually based on comparisons of
striae in land engraved areas, which are extracted in form of cross
sections, called \emph{profiles} \citep{aoas,ma2004}. Bullet striae are
most pronounced at the bottom of the bullet heel (because the heel has
the most contact with the insides of the barrel). Howver, these areas
are also affected by the break off due to friction effects between
barrel and the bullet. An optimal cross section is chosen orthogonally
to the striae, close to the heel while avoiding break off as shown in
\autoref{fig:rgl}, see also \citep{aoas} for mathematical details of the
extraction.

From profiles bullet \emph{signatures} \citep{chu2013,aoas} are
extracted as residuals of a loess fit or Gaussian filter. This
effectively removes topographic structure from the data in the attempt
to increase the signal to noise ratio. \autoref{fig:rgl} shows how the
signature from a bullet land (bottom) lines up with the image of the
land (top) from which it was extracted. We can see in the figure how the
depth and relative position of the striation markings seen in the image
are interpreted as peaks and valleys in the signature.

@@

For both same- and different-shift pairs correlations between the
markings are calculated. The intuition here is that for two markings
from the same-source the correlation for the in-phase sample should be
high, while the correlations of the out-of-phase sample provide a
measure for the base-level correlation for non-matching marks of a given
length \(w_v\).
{\textcolor{orange}{More specifically, the null hypothesis of the Chumbley score test is stated as **$H_o$: the markings come from two different sources**}}
with the alternative given as \textbf{\(H_a\): the markings come from
the same source}. A \(p\)-value for the Chumbley score test is \} then
computed as a Mann Whitney U statistic to compare between in-phase
sample and out-of-phase sample. A low \(p\)-value is then interpreted as
a rejection of the null hypothesis in favor of the alternative
hypothesis of \texttt{same-source}.\}

In the original method proposed in \citet{chumbley} both in-phase and
out-of-phase sample are extracted randomly, whereas \citet{hadler}
proposed the above specified deterministic rules for both samples to
make the resulting score deterministic while simultaneously avoiding
overlaps within selected marks to ensure independence.

@@

Looking closer at \autoref{eq.start}, we see that by definition, some
number of tests will fail to produce a result.
{\textcolor{orange}{Note that this problem is different from erroneous test results.}}
The problem of failed tests is first mentioned in \citet{afte-chumbley}.
Unfortunately, the authors do not provide any percentage of how many
tests failed for their data. The algorithm fails to produce for two
reasons: either the number of eligible same-shift pairs is zero, or the
number of different-shift pairs is zero.

@@

We then define the pairs for different-shifts as
\begin{equation}\label{eq.diff2}
\left(s_{(j)}^{(1)}, s_{(I-j+1)}^{(1)} \right) \text{ for } j = 
\begin{cases}
1, ..., I & \text{ for even } I \\
1, ..., (I-1)/2, (I-1)/2 + 2, ..., I & \text{ for odd } I
\end{cases},
\end{equation} i.e.~for an odd number of same-shift correlations, we
skip the middle pair for the different-shift correlations (see
\autoref{sketch-diff-2}). This pairing ensures that the number of
different-shift pairings is the same or at most one less than the number
of same-shift pairings in all tests. In the remainder of the paper, we
will refer to the algorithm defined by \citet{hadler} as \textbf{(CS1)}
and the suggested modified algorithm as \textbf{(CS2)} and compare their
performance on the available scans of the Hamby study.

{\textcolor{orange}{Performance of tests is measured with respect to the errors they make in situations where ground truth is known. We distinguish between two error rates: (1) false positive and (2) false negative rate.
*False positives* (or *false identifications*) are situations where the test indicates a match (i.e. the test falsely rejects) but the markings come from different sources. This is also known as a *Type I* error. 
*False negatives* (or *missed identifications*) are situations where the test fails to reject, i.e. the test indicates that the markings come from different sources, but in fact the markings come from the same source. This is a *Type II* error.}}

{\textcolor{orange}{Note that in all of the following land-to-land comparisons only lands are compared that are suitable for a comparison, i.e. a signature can be extracted from the scan. In particular, lands which exhibited tank rash  were removed from comparison \citep{aoas}.
}}

The section \ref{appendix:appxfailed} of the Appendix discusses about
the various scenarious of failed tests.

@@

\begin{table}[hbtp]
\centering
\begin{adjustbox}{max width=\textwidth}
\begin{tabular}{rrrlrrl}
  \hline
  & \multicolumn{3}{l}{\bf CS1} & \multicolumn{3}{l}{\bf CS2} \\
$w_o$ & \bf FPR & \bf FNR & \bf AUC (95\% C.I.) & \bf FPR & \bf FNR & \bf AUC (95\% C.I.) \\  
  \hline
90 & 0.066 & 0.512 & 0.680 (0.633, 0.727) & 0.066 & 0.512 & 0.729 (0.684, 0.774) \\ 
  120 & 0.069 & 0.468 & 0.736 (0.693, 0.779) & 0.069 & 0.468 & 0.763 (0.720, 0.806) \\ 
  140 & 0.070 & 0.473 & 0.741 (0.698, 0.784) & 0.070 & 0.473 & 0.767 (0.727, 0.808) \\ 
  150 & 0.067 & 0.457 & 0.730 (0.685, 0.775) & 0.067 & 0.457 & 0.774 (0.733, 0.816) \\ 
  180 & 0.071 & 0.446 & 0.773 (0.734, 0.813) & 0.071 & 0.446 & 0.773 (0.730, 0.816) \\ 
  210 & 0.070 & 0.400 & 0.780 (0.740, 0.821) & 0.070 & 0.400 & 0.803 (0.763, 0.843) \\ 
  240 & 0.070 & 0.396 & 0.784 (0.744, 0.824) & 0.070 & 0.396 & 0.801 (0.760, 0.841) \\ 
  270 & 0.065 & 0.394 & 0.782 (0.741, 0.823) & 0.065 & 0.394 & 0.808 (0.769, 0.848) \\ 
  300 & 0.066 & 0.394 & 0.791 (0.751, 0.832) & 0.066 & 0.394 & 0.814 (0.776, 0.852) \\ 
  330 & 0.066 & 0.405 & 0.791 (0.753, 0.830) & 0.066 & 0.405 & 0.804 (0.765, 0.844) \\ 
  360 & 0.066 & 0.413 & 0.788 (0.747, 0.828) & 0.066 & 0.413 & 0.803 (0.765, 0.842) \\ 
  390 & 0.067 & 0.391 & 0.789 (0.748, 0.830) & 0.067 & 0.391 & 0.810 (0.771, 0.849) \\ 
  480 & 0.063 & 0.419 & 0.755 (0.711, 0.800) & 0.063 & 0.419 & 0.801 (0.761, 0.841) \\ 
   \hline
\end{tabular}
\end{adjustbox}
\caption{\label{tab:aucsh44} Overview of high resolution scan results as shown in Figures \ref{fig:h44}. FPR is the observed rate of false positives (for a fixed level $\alpha = 0.05$), FNR is the false negative rate. Area under the curve (AUC) is shown with confidence intervals. }
\end{table}
